---
title: "Scrap Reference"
author: "Aaron Simot"
date: "10/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Loading various libraries
```{r loads, include=FALSE}
library(rvest)
library(tidyverse)
```



Below is the test to extract a table from the defined URL, but using the team standard batting reference
```{r dataframe_set, message=FALSE, include = FALSE}
newurl <- read_html("https://www.baseball-reference.com/leagues/majors/2020-standard-batting.shtml")

#Team Standard Batting 2020
tester <- newurl %>%
  #retrieve elements with table in HTML
  html_elements('table') %>%
  #create a table from HTML
  html_table()

tail(tester)

```


Define a function to automate extracts for team standard batting
```{r auto_extract_team, message=FALSE}

#input parameter _x_ is a URL to be scraped
url_reads <- function(x) {
  #declare result as returned dataframe of team batting standards 
  result <- x %>% 
    read_html() %>% #read the url string to HTML
    html_elements('table') %>% #read the elements with reference table
    html_table() #create table
  
  #result was saved as a list with dataframe at index1, pipe that into mutate
  result <- result[[1]] %>%
    #adding a year column based on the input URL and extract the year from the URL with regex and sending char to numeric
    #casting all columns to be characters.  Had merging frames issue where column types did not agree - Aaron 10/22
    mutate(year = as.numeric(str_sub(x, start = str_locate(x, "\\d\\d\\d\\d"), end = str_locate(x, "standard"))),
           across(everything(), as.character)) 
  
  result #return the result dataframe from URL _x_
}

# Tried this and was not working properly -> something wrong with the filter before for loop
# url_reads <- function(x) {
#   #initial load frame -- need a frame to start to bind to 
#   result <- x[1] %>% 
#     read_html() %>% 
#     html_elements('table') %>%
#     html_table()
#   
#   #start at index two since we initialized on index one
#   result2 <- result %>% filter(is.na(Tm))
#   for (i in 2:length(x)) {
#     #loop over array of urls for each each year 
#     result2 <- x[i] %>%
#       read_html() %>%
#       html_elements('table') %>%
#       html_table() %>%
#     
#     result <- rbind(result, result2) #bind results to initial frame
#     
#   }
# }

```

Code chunk to create vector of URLs to scrape based on years vector.  Applies map_df to our URL vector and the url_reads function to create a comprehensive list of team batting standards from 1876 to 2020
```{r define_df, message = FALSE}
#vector of all record years to scrape
#there is an issue with a year 1960-1964 where #Bat is char not numeric causing an issue in combining dataframes
#Look into this later 10/21 - Aaron [issue fixed 10/22 using char as all column types]
years <- as.character(c(1900:2020))

#vector of years to test design, subset of years vector
#no longer needed, testing is completed
#test_years <- as.character(c(2010:2020))

#list of URLS for Team Standard Batting tables
team_urls_bat <- str_c("https://www.baseball-reference.com/leagues/majors/",years,"-standard-batting.shtml")

#create new dataframe of all years extracted using map_df to specify vector or URLs and function (url_reads) to apply
teams_batting <- map_df(team_urls_bat, url_reads)
# Around 40 second execution time for years 1900-2020

```

```{r export_data}

#export the team batting standard dataframe to a csv
write_csv2(teams_batting, here::here('teamsBattingStandards1900_2020'))

```

Code chunk for team pitching data
```{R pitching_data}

#array for URLs of teams pitching standards 1900-2020
teams_urls_pit <- str_c("https://www.baseball-reference.com/leagues/majors/",years,"-standard-pitching.shtml")

#mapping the teams pitching standard dataframe
teams_pitching <- map_df(teams_urls_pit, url_reads)

```


```{R extract_pitch}
#extracting teams pitching standards
write_csv2(teams_pitching, here::here('teamsPitchingStandards1900_2020'))

```



Frame to test url_reads function defined earlier
```{r testing_returned_results, include = FALSE}
#test URL
y <- "https://www.baseball-reference.com/leagues/majors/2020-standard-batting.shtml"

#Testing function call
test2020 <- url_reads(y)
#return result
test2020
```





